---
title: "ADS 503: Final Team Project"
author: "Caleb McCurdy, Mirna Philip, and Christine Vu"
date: "June 26, 2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

## Data Importing and Pre-processing
```{r message = FALSE, warning = FALSE, results = 'hide'}
library(readr)
library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)
library(corrplot)
library(mlbench)
library(moments)
library(e1071)
library(caret)
library(Hmisc)
library(plotly)
library(psych)
library(MASS)
library(AppliedPredictiveModeling)
library(glmnet)
```

```{r}
df <- read.table('https://raw.githubusercontent.com/mirna1983/dermatology-project/main/data/dermatology.data', sep = ",")

head(df, 6)
```

```{r}
# Dimension of data (rows and columns)
dim(df)
```

&nbsp;

#### Add Column Names

The data frame is missing the column names that correspond to the provided attribute list. Therefore, the column names are appended based on the attribute information list.

```{r, warnings = FALSE}
# Desired column names
desired_columns <- c("erythema", "scaling", "definite_borders", "itching", "koebner_phenomenon", "polygonal_papules", "follicular_papules", "oral_mucosal_involvement", "knee_and_elbow_involvement", "scalp_involvement", "family_history", "melanin_incontinence", "eosinophils_in_the_infiltrate", "PNL_infiltrate", "fibrosis_of_the_papillary_dermis", "exocytosis", "acanthosis", "hyperkeratosis", "parakeratosis", "clubbing_of_the_rete_ridges", "elongation_of_the_rete_ridges", "thinning_of_the_suprapapillary_epidermis", "spongiform_pustule", "munro_microabcess", "focal_hypergranulosis", "disappearance_of_the_granular_layer", "vacuolisation_and_damage_of_basal_layer", "spongiosis", "saw-tooth_appearance_of_retes", "follicular_horn_plug", "perifollicular_parakeratosis", "inflammatory_monoluclear_inflitrate", "band-like_infiltrate", "Age", "class")

# Assign desired column names
colnames(df) <- desired_columns
head(df, 3)
```

```{r}
# Summary of the data frame
summary(df)
```

```{r}
str(df)
```

```{r}
# Change 'Age' into type numeric
df$Age <- as.numeric(df$Age)
# Change 'Class' into type factor
df$class <- as.factor(df$class)
```

```{r}
# Check for missing values
sapply(df, function(x) sum(is.na(x)))
```

There are no missing values in this data set other than for the 'Age' predictor which has 8. Due to the small number of instances with missing values, we will delete these rows from the data set.

```{r}
rows_with_NA <- which(is.na(df$Age))
rows_with_NA

df <- df[-rows_with_NA, ]
head(df, 6)
```

```{r}
# Check for potential duplicate instances
sum(duplicated(as.matrix(df), fromLast = FALSE))
```

There are no duplicate instances to deal with.

```{r}
# Make the graphs smaller by having 3 rows of graphs and 4 columns 
par(mfrow = c(3,4))

hist.data.frame(df)
```

```{r}
ggplot(df, aes(class, Age)) +
geom_boxplot(colour = "blue", outlier.colour = "red") + ggtitle("Erythemato-Squamous Disease based on Age") + xlab("Classification of Disease") + ylab("Age of the patient in Years") +
theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(df, aes(class, fibrosis_of_the_papillary_dermis)) +
geom_boxplot(colour = "Blue", outlier.colour = "red") +
ggtitle("Erythemato-Squamous Disease based on Papillary Dermis Fibrosis") + xlab("ESD Classification") +
ylab("Fibrosis of Papillary Dermis") +
theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(df, aes(class, thinning_of_the_suprapapillary_epidermis)) +
geom_boxplot(colour = "Blue", outlier.colour = "red") +
ggtitle("Erythemato-Squamous Disease based on Suprapapillary Epidermis Thinning") + xlab("ESD Classification") +
ylab("Thinning of Suprapapillary Epidermis") +
theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(df, aes(class, perifollicular_parakeratosis)) +
geom_boxplot(colour = "Blue", outlier.colour = "red") +
ggtitle("Erythemato-Squamous Disease based on Perifollicular Parakeratosis") + xlab("ESD Classification") +
ylab("Perifollicular Parakeratosis Level") +
theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(df, aes(class, clubbing_of_the_rete_ridges)) +
geom_boxplot(colour = "Blue", outlier.colour = "red") +
ggtitle("Erythemato-Squamous Disease based on Rete Ridges Clubbing") + xlab("ESD Classification") +
ylab("Clubbing of Rete Ridges") +
theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(df, aes(class, itching)) +
geom_boxplot(colour = "Blue", outlier.colour = "red") +
ggtitle("Erythemato-Squamous Disease based on Itching") + xlab("ESD Classification") +
ylab("Itching Severity") +
theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# Plotting the correlation matrix using the corrplot function, only using 
# complete pairs of observations between variables
corrplot(cor(df[, -35]), cl.cex= .5,tl.cex=.4)
```

```{r}
# plotting boxplots of the 34 predictors and response variable
boxplot(df[1:7])
boxplot(df[8:14])
boxplot(df[15:21])
boxplot(df[22:28])
boxplot(df[29:35])
```

```{r}
# Skewness of numeric predictors
skewness_values <- apply(df[, 1:34], 2, skewness, na.rm = TRUE)
skewness_values
```

```{r}
pred <- df[,-35]
class <- df$class

levels(class) <- c("A", "B", "C", "D", "E", "F")

# Split the data into training (80%) and test (20%) sets
set.seed(123)
training <- createDataPartition(class, p = 0.8, list = FALSE)
pred_train <- pred[training,]
class_train <- class[training]
pred_test <- pred[-training,]
class_test <- class[-training]

# pre-process the data
preProcValues <- preProcess(pred_train, method = c("center", "scale"))

pred_train_trans <- predict(preProcValues, pred_train)
pred_test_trans <- predict(preProcValues, pred_test)

ctrl <- trainControl(method = "cv",
                     summaryFunction = multiClassSummary,
                     classProbs = TRUE,
                     savePredictions = TRUE)
```

#### Neural networks
```{r, warning = FALSE}
nnetGrid <- expand.grid(decay = c(0, 0.01, .1), size = c(3, 7, 11, 13))

set.seed(123)
nnetTune <- train(x = pred_train_trans, y = class_train,
                  method = "nnet",
                  tuneGrid = nnetGrid,
                  trControl = ctrl,
                  linout = TRUE,
                  trace = FALSE,
                  MaxNWts = 13 * (ncol(pred_train_trans) + 1) + 13 + 1,
                  maxit = 1000)

nnet_pred <- predict(nnetTune, pred_test_trans)

confusionMatrix(nnet_pred, class_test)
```


#### k-nearest neighbors
```{r, warning = FALSE}
knnDescr <- pred_train_trans[, -nearZeroVar(pred_train_trans)]

set.seed(123)
knnTune <- train(x = knnDescr, y = class_train,
                 method = "knn",
                 tuneGrid = data.frame(k = 1:20),
                 trControl = ctrl)

knn_pred <- predict(knnTune, pred_test_trans[, names(knnDescr)])

confusionMatrix(knn_pred, class_test)
```


#### Support vector machine (Linear)
```{r, warning = FALSE}
set.seed(123)
svm_lin <- train(x = pred_train_trans, y = class_train,
                  method = "svmLinear", tuneLength = 10,
                  trControl = ctrl)

svml_pred <- predict(svm_lin, pred_test_trans)

confusionMatrix(svml_pred, class_test)
```


#### Support vector machine (RBF)
```{r, warning = FALSE}
set.seed(123)
svm_rad <- train(x = pred_train_trans, y = class_train,
                  method = "svmRadial", tuneLength = 10,
                  trControl = ctrl)

svmr_pred <- predict(svm_rad, pred_test_trans)

confusionMatrix(svmr_pred, class_test)
```


#### Linear Discriminant Analysis
```{r}
set.seed(123)
ldaFit <- train(x = pred_train_trans,
                y = class_train,
                method = "lda",
                metric = "Accuracy",
                trControl = ctrl)

lda_pred <- predict(ldaFit, pred_test_trans)

confusionMatrix(lda_pred, class_test)
```


#### Penalized Logistic Regression
```{r, warning = FALSE}
glmnGrid <- expand.grid(alpha = c(0,  .1,  .2, .4, .6, .8, 1),
                        lambda = seq(.01, .2, length = 10))

glmnFit <- train(x = pred_train_trans,
                 y = class_train,
                 method = "glmnet",
                 tuneGrid = glmnGrid,
                 metric = "Accuracy",
                 trControl = ctrl)

glm_pred <- predict(glmnFit, pred_test_trans)

confusionMatrix(glm_pred, class_test)
```


#### Nearest Shrunken Centroids
```{r, warning = FALSE}
nscFit <- train(x = pred_train_trans,
                y = class_train,
                method = "pam",
                tuneGrid = data.frame(threshold = seq(0, 25, length = 30)),
                metric = "Accuracy",
                trControl = ctrl)

nsc_pred <- predict(nscFit, pred_test_trans)

confusionMatrix(nsc_pred, class_test)
```



```{r}
# Variable importance
nnetImp <- varImp(nnetTune, scale = FALSE)
#plot(nnetImp, top = 5)
knnImp <- varImp(knnTune, scale = FALSE)
plot(knnImp, top = 5)
svmLImp <- varImp(svm_lin, scale = FALSE)
plot(svmLImp, top = 5)
svmRImp <- varImp(svm_rad, scale = FALSE)
plot(svmRImp, top = 5)
ldaImp <- varImp(ldaFit, scale = FALSE)
plot(ldaImp, top = 5)
glmImp <- varImp(glmnFit, scale = FALSE)
plot(glmImp, top = 5)
nscImp <- varImp(nscFit, scale = FALSE)
plot(nscImp, top = 5)
```

